When Do Repeat Yourself?\\ Voices from the Trenches of Linux Kernel Maintainers on Code Duplication

### 2025-10-24 paper/techdebt2026/techdebt.tex
techdebt2026: update title
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

\title{Repeat Yourself? When Linux kernel maintainers balance Code Duplication and Readability}

@@

In the AMD Display driver codebase, we executed our tool. We manually analyzed the biggest duplications regarding the number of lines and found two duplicated function pairs that we judged interesting to try to mitigate. We have chosen to approach the biggest duplications as proof of concept so that we can make significant impacts.

Given the two duplicated function pairs, we observed that the files in the functions 
contain multiple other duplications. Thus, we proposed a simple systematic approach 
to mitigate all the functions duplicated in the context, not just the duplicated 
function pairs. 
After we approached the duplications with the systematic strategy, for the duplicated 
function pairs that had successfully mitigated the duplication, we sent patches to the 
AMD Display driver email list to validate the mitigations and see the maintainer's opinion 
while documenting the process and our impressions.
The systematic approach is defined below.

@@

In the \textbf{complete participant observation} study, we chose two function pair found on AMD Display Driver to approach and apply the systematic approach ...

@@

Firt, we found that the duplicated function from mitigation 1
exists in 10 code files, enclosing multiple GPU architectures. The duplicated 
functions are not exactly equal, as some minimal specific logic is applied to each GPU architecture. To address this issue, we used the extract method to 
break the function into smaller parts and the parameterized method to address 
the specific logic for the GPUs.

We reached a point where we had a scratch on how to mitigate the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver. We found that the configuration files on the AMD Display driver are designed to be imported at compilation time using \textit{\#define} macros. The configuration files' design choice makes refactoring duplications on generic approaches trickier, as refactoring code that depends on these files requires significant refactoring of the configuration files' design and deep knowledge of the codebase, which we, as first-time contributors, do not have. Thus, we opted not to continue investigating this mitigation.

Second, we found that the duplicated function from mitigation 2 exists in five code files, 
and two other functions are duplicated in those five code files. All functions across the files were exactly equal, 
so there was no need to apply the refactoring methods presented in the literature. 
The refactoring was resumed to create a generic library and fix the compilation targets.

The code files in the second mitigation do not depend on configuration files. Thus, 
we did not face the same issues during the first mitigation. Refactoring the code 
to mitigate duplications reached a point we judged was good enough to submit as a 
patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
the refactoring to the driver maintainers as a patch, documenting our findings and experiences 
on the process.

@@

%TODO: deve ficar claro que isto aqui, mesmo que não dito explicitamente pelo mantenedores, é o resultado da Complete Participant Observation
\noindent\textbf{Takeaway T1:} Some of the code in the driver is shared across all operating systems that support the implemented GPUs. This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors. Additionally, \textbf{we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering}. One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others. This approach helps save a significant amount of time and effort.
\todo[inline]{...Foi tirado apenas das conversas com o Siqueira; tem o paper que o Gregório citou que menciona que duplicações podem ser bem vistas em drivers; isto um resultado/conclusão da Complete Participant Observation; e vamos citar que confirmamos o que está no paper Cloning Considered Harmful (Luan, acha que consegue fechar essa parte assim?)}

>>>>

\title{When Do Repeat Yourself?\\
Voices from the Trenches of Linux Kernel Maintainers on Code Duplication}

@@

In the AMD Display driver codebase, we executed our tool. We manually analyzed the biggest duplications regarding the number of lines and found one duplicated function pairs that we judged interesting to try to reduce the duplication. We have chosen to approach the biggest duplications as proof of concept so that we can make significant impacts.

Given the duplicated function pair, we observed that the files in the functions
contain multiple other duplications. Thus, we proposed a simple systematic approach 
to mitigate all the functions duplicated in the context, not just the duplicated 
function pairs. 
After we approached the duplications with the systematic strategy, we sent patch to the 
AMD Display driver email list to validate the mitigations and see the maintainer's opinion 
while documenting the process and our impressions.
The systematic approach is defined below.

@@

In the \textbf{complete participant observation} study, we chose one function pair found on AMD Display Driver to approach and apply the systematic approach.

@@

We reached a point where we had a scratch on how to mitigate the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver. We found that the configuration files on the AMD Display driver are designed to be imported at compilation time using \textit{\#define} macros. The configuration files' design choice makes refactoring duplications on generic approaches trickier, as refactoring code that depends on these files requires significant refactoring of the configuration files' design and deep knowledge of the codebase, which we, as first-time contributors, do not have. Thus, we opted not to continue investigating this mitigation.
We found that the duplicated function from the code redution exists in five code files, 
and two other functions are duplicated in those five code files. All functions across the files were exactly equal, 
so there was no need to apply the refactoring methods presented in the literature. 
The refactoring was resumed to create a generic library and fix the compilation targets.

Refactoring the code to reduce the duplications reached a point we judged was good enough to submit as a 
patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
the refactoring to the driver maintainers as a patch, documenting our findings and experiences 
on the process.

@@

%TODO: deve ficar claro que isto aqui, mesmo que não dito explicitamente pelo mantenedores, é o resultado da Complete Participant Observation
\noindent\textbf{Takeaway T1:}
\textbf{We came to understand that AMD Display driver developers do not view duplicated code negatively, 
which seems to contradict standard software engineering practices.} 
In our attempt to understand the reason for the delay in the review process, 
we contacted a maintainer. It was mentioned to us that one reason for the delay 
is that duplicated code enhances the independence of GPU driver code. This allows 
developers to make changes to a specific GPU without needing to test compatibility 
with others, which helps save a significant amount of time and effort. This example of 
hardware variations, also mentioned by Kapser and Godfrey~\cite{cloneharm}, 
corroborates our understanding that driver developers do not view duplicated code negatively by default.
